---
title: "Simulating Data"
output: html_document
date: "2023-07-11"
---

```{r}
setwd(paste("C:/Users/ezequ/OneDrive - Fundacao Getulio Vargas - FGV/",
            "Grad MAp FGV/IC/modeling-brazilian-soccer", sep = ""))

source("elo_rating/simulation/code/functions.R")
```

# Model

Given two teams $A$ and $B$ with ratings $r_A$ and $r_B$, the model is

```{=latex}
\begin{align*}
  \mathbb{P}(A \ wins) = \frac{r_A^2}{r_A^2 + r_B^2 + r_A r_B k}
  \\
  \mathbb{P}(B \ wins) = \frac{r_B^2}{r_A^2 + r_B^2 + r_A r_B k}
  \\
  \mathbb{P}(draw) = \frac{r_A r_B k}{r_A^2 + r_B^2 + r_A r_B k}

\end{align*}
```
for $k > 0$.

# Simulation without estimating k

Given *n_teams* teams, *n_matches* matches between teams (home and away), a mean
*mu* and a standard deviation *sd*, the *sim_data()* function generates ratings 
from normal distribution for teams and simulates the results for matches.

### Two teams ($k = 0$)

For small number of matches, the result is bad:

```{r}
n_teams <- 2
n_matches <- 50
mu <- 20
sd <- 4
k <- 0

sim_data(n_teams, n_matches, mu, sd, k)

rating <- read.csv("elo_rating/simulation/data/simulated_rating.csv")
results <- read.csv("elo_rating/simulation/data/simulated_results.csv")

initial_values <- c(rep(100, n_teams-1))
opt <- optim(initial_values, fn = lfn_elo_rating, results = results,
             lower= rep(10^(-10), n_teams-1), method="L-BFGS-B")
rating$r_hat <- c(1, opt$par)

rating
```

For a large number, the result improves:

```{r}
n_teams <- 2
n_matches <- 1000
mu <- 20
sd <- 4
k <- 0

sim_data(n_teams, n_matches, mu, sd, k)

rating <- read.csv("elo_rating/simulation/data/simulated_rating.csv")
results <- read.csv("elo_rating/simulation/data/simulated_results.csv")

initial_values <- c(rep(100, n_teams-1))
opt <- optim(initial_values, fn = lfn_elo_rating, results = results,
             lower= rep(10^(-10), n_teams-1), method="L-BFGS-B")
rating$r_hat <- c(1, opt$par)

rating
```

### Two teams ($k > 0$)

```{r}
n_teams <- 2
n_matches <- 1000
mu <- 20
sd <- 4
k <- 5

sim_data(n_teams, n_matches, mu, sd, k)

rating <- read.csv("elo_rating/simulation/data/simulated_rating.csv")
results <- read.csv("elo_rating/simulation/data/simulated_results.csv")

initial_values <- c(rep(100, n_teams-1))
opt <- optim(initial_values, fn = lfn_elo_rating_with_draw_k, results = results, 
             k = k, lower= rep(10^(-10), n_teams-1), method="L-BFGS-B")
rating$r_hat <- c(1, opt$par)

rating
```

### More than two teams ($k = 0$)

```{r}
n_teams <- 5
n_matches <- 20
mu <- 20
sd <- 4
k <- 0

sim_data(n_teams, n_matches, mu, sd, k)

rating <- read.csv("elo_rating/simulation/data/simulated_rating.csv")
results <- read.csv("elo_rating/simulation/data/simulated_results.csv")

initial_values <- c(rep(100, n_teams-1))
opt <- optim(initial_values, fn = lfn_elo_rating, results = results,
               lower= rep(10^(-10), n_teams-1), method="L-BFGS-B")
rating$r_hat <- c(1, opt$par)

rating

```

### More than two teams ($k > 0$)

```{r}
n_teams <- 5
n_matches <- 20
mu <- 20
sd <- 4
k <- 5

sim_data(n_teams, n_matches, mu, sd, k)

rating <- read.csv("elo_rating/simulation/data/simulated_rating.csv")
results <- read.csv("elo_rating/simulation/data/simulated_results.csv")

initial_values <- c(rep(100, n_teams-1))
opt <- optim(initial_values, fn = lfn_elo_rating_with_draw_k, results = results, 
             k = k, lower= rep(10^(-10), n_teams-1), method="L-BFGS-B")
rating$r_hat <- c(1, opt$par)

rating

```

# Simulation estimating k

### Two teams

```{r}
n_teams <- 2
n_matches <- 1000
mu <- 20
sd <- 4
k <- 5

sim_data(n_teams, n_matches, mu, sd, k)

rating <- read.csv("elo_rating/simulation/data/simulated_rating.csv")
results <- read.csv("elo_rating/simulation/data/simulated_results.csv")

initial_values <- c(rep(100, n_teams))
opt <- optim(initial_values, fn = lfn_elo_rating_with_draw, results = results,
             lower= rep(10^(-10), n_teams), method="L-BFGS-B")
rating$r_hat <- c(1, opt$par[1:(n_teams-1)])
df_k <- data.frame(k = k,
                   k_hat = opt$par[n_teams])

rating
df_k
```

### More than two teams

```{r}
n_teams <- 5
n_matches <- 20
mu <- 20
sd <- 4
k <- 8

sim_data(n_teams, n_matches, mu, sd, k)

rating <- read.csv("elo_rating/simulation/data/simulated_rating.csv")
results <- read.csv("elo_rating/simulation/data/simulated_results.csv")

initial_values <- c(rep(100, n_teams))
opt <- optim(initial_values, fn = lfn_elo_rating_with_draw, results = results,
             lower= rep(10^(-10), n_teams), method="L-BFGS-B")
rating$r_hat <- c(1, opt$par[1:(n_teams-1)])
df_k <- data.frame(k = k,
                   k_hat = opt$par[n_teams])

rating
df_k
```

